Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-04-20 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update MS

{\it POPC with 128 lipids}
The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website (http://people.su.se/$\sim$jjm/Stockholm\_Lipids/Downloads.html).
The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
They contained 100 POPC/24 cholesterol molecules and 80 POPC/80 cholesterol molecules for
the simulations of 20\% cholesterol and 50\% cholesterol respectively. The TIP3P water model~\cite{jorgensen83} was used to
solvate the system.
The publicly available CHARMM36 forcefield parameters (\url{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool~\cite{salari15}.  
Single point energy calculation was done to assess the conversion. 
%(At the time I've made the simulations, the charmm36 ff for cholesterol in gromacs format was not available at
%http://mackerell.umaryland.edu/charmm_ff.shtml#gromacs).
Simulations were performed for 200~ns and the last 100~ns was used for the calculations. Timestep of 2~fs was
used with leapfrog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
303~K with the velocity-rescale method~\cite{bussi07} and a time constant of 0.2~ps. Pressure was maintained semiisotropically
at 1~bar using the Parrinello--Rahman algorithm~\cite{parrinello81} with a time constant of 1.0~ps. The neighbour list with a cut-off of 1.2~nm
was updated every 10 steps. Lennard-Jones interactions were switched to zero
between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics.

@@

\subsubsection{Slipids}
Initial coordinates for a hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the 
Slipids home page \url{http://people.su.se/\~jjm/Stockholm\_Lipids/Downloads.html}.  The Slipids force field~\cite{jambeck12,jambeck12b} was used for the the all atom descriptions of DPPC and POPC, and
water was described with the TIP3P water model~\cite{jorgensen83}. Simulations were performed within the NPT ensemble using the GROMACS 4.6.X simulation
package~\cite{hess08}. The Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} was used with reference temperatures of
323~K (DPPC) and 310~K (POPC) and a relaxation time constant of 0.5~ps. Water and lipids were coupled separately to
the heat bath. Pressure was kept constant at 1.013~bar using a semi--isotropic Parrinello--Rahman
barostat~\cite{parrinello81} with a time constant of 10.0~ps. Equations of motion were
integrated with the leapfrog algorithm using a timestep of 2~fs. Long range
electrostatic interactions were calculated using the PME method~\cite{darden93,essman95}, with a fourth order
smoothing spline. A real space cut-off of 1.0~nm was employed with grid spacing of 0.12~nm in the reciprocal space.
Lennard-Jones potentials were cut off at 1.4~nm, with a dispersion correction applied to both energy and pressure. All covalent bonds in lipids were constrained using the LINCS algorithm~\cite{hess97}, 
whereas water molecules were constrained using SETTLE~\cite{miyamoto92}. Twin-range cutoffs,
1.0~nm and 1.6~nm, were used for the neighbor lists with the longrange neighbor list updated every
10 steps. This simulation protocol corresponds to the protocol used in Ref~\cite{jambeck13}.

>>>>

{\it POPC with 128 lipids}
The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website (\url{http://mmkluster.fos.su.se/slipids/Downloads.html}).
The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
They contained 100 POPC/24 cholesterol molecules and 80 POPC/80 cholesterol molecules for
the simulations of 20\% cholesterol and 50\% cholesterol respectively. The TIP3P water model~\cite{jorgensen83} was used to
solvate the system.
The publicly available CHARMM36 forcefield parameters (\url{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool~\cite{salari15}.  
Single point energy calculation was done to assess the conversion. 
%(At the time I've made the simulations, the charmm36 ff for cholesterol in gromacs format was not available at
%http://mackerell.umaryland.edu/charmm_ff.shtml#gromacs).
Simulations were performed for 200~ns and the last 100~ns was used for the calculations. Timestep of 2~fs was
used with leapfrog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
303~K with the velocity-rescale method~\cite{bussi07} and a time constant of 0.2~ps. Pressure was maintained semiisotropically
at 1~bar using the Parrinello--Rahman algorithm~\cite{parrinello81} with a time constant of 1.0~ps. The neighbour list with a cut-off of 1.2~nm
was updated every 10 steps. Lennard-Jones interactions were switched to zero
between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics.

@@

\subsubsection{Slipids}
Initial coordinates for a hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the 
Slipids home page \url{http://mmkluster.fos.su.se/slipids/Downloads.html}.  The Slipids force field~\cite{jambeck12,jambeck12b} was used for the the all atom descriptions of DPPC and POPC, and
water was described with the TIP3P water model~\cite{jorgensen83}. Simulations were performed within the NPT ensemble using the GROMACS 4.6.X simulation
package~\cite{hess08}. The Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} was used with reference temperatures of
323~K (DPPC) and 310~K (POPC) and a relaxation time constant of 0.5~ps. Water and lipids were coupled separately to
the heat bath. Pressure was kept constant at 1.013~bar using a semi--isotropic Parrinello--Rahman
barostat~\cite{parrinello81} with a time constant of 10.0~ps. Equations of motion were
integrated with the leapfrog algorithm using a timestep of 2~fs. Long range
electrostatic interactions were calculated using the PME method~\cite{darden93,essman95}, with a fourth order
smoothing spline. A real space cut-off of 1.0~nm was employed with grid spacing of 0.12~nm in the reciprocal space.
Lennard-Jones potentials were cut off at 1.4~nm, with a dispersion correction applied to both energy and pressure. All covalent bonds in lipids were constrained using the LINCS algorithm~\cite{hess97}, 
whereas water molecules were constrained using SETTLE~\cite{miyamoto92}. Twin-range cutoffs,
1.0~nm and 1.6~nm, were used for the neighbor lists with the longrange neighbor list updated every
10 steps. This simulation protocol corresponds to the protocol used in Ref~\cite{jambeck13}.
